---
title: "Quantity judgements in stingless bees"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(tidyboot)
library(lme4)
library(ggthemes)
library(ggridges)
library(lsr)
library(broom.mixed)
library(emmeans)

```

# Data

```{r}
# read_xlsx("../data/stinglesbees_rawdata.xlsx", sheet = 3)%>%
#   select(-time)%>%
#   write_csv("../data/data.csv")

data <- read_csv("../data/data.csv")%>%
  mutate(subject = as.factor(subject))

```

# Sample size

```{r}
data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface")))%>%
  group_by(experiment)%>%
  summarise(N = length(unique(subject)))
```

# Visualize data by experiment

```{r}
p1 <- data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface"), labels  = c("Quantity only", "Quantity vs. shape", "Quantity vs. color", "Quantity vs. surface area")))%>%
  group_by(subject,experiment)%>%
  summarise(mean = mean(correct))#

p2 <- data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface"), labels  = c("Quantity only", "Quantity vs. shape", "Quantity vs. color", "Quantity vs. surface area")))%>%
  group_by(experiment)%>%
  tidyboot_mean(col = correct)

ggplot()+
  geom_hline(yintercept = 0.5, lty = 2, alpha = .5)+
  geom_violin(data = p1, aes(x = experiment, y = mean), alpha = .25, fill = "grey")+
  geom_point(data = p1, aes(x = experiment, y = mean), alpha = .5, col = "darkgrey")+
  geom_pointrange(data = p2, aes(x = experiment, y = mean, ymin = ci_lower, ymax = ci_upper), col = "firebrick")+
  ylim(0,1)+
  labs(y = "Proportion of choices based on quantity", x = "Experiment")+
  guides(col = F, fill = F)+
  scale_color_colorblind()+
  theme_minimal()
    
```
# Visualize data by experiment, sample quantity and feature (and overall)

```{r}

p2 <- p2 %>%
  mutate(sample_quantity = "overall",
         feature = "-")

p3 <- data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface"), labels  = c("Quantity only", "Quantity vs. shape", "Quantity vs. color", "Quantity vs. surface area")))%>%
  mutate(sample_quantity = factor(sample_quantity),
         feature = ifelse(experiment == "Quantity vs. surface area", feature_of_numerical_match, sample_feature),
         feature = ifelse(experiment == "Quantity only", "-",feature))%>%
  group_by(subject,sample_quantity,feature, experiment)%>%
  summarise(mean = mean(correct))#


p4 <- data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface"), labels  = c("Quantity only", "Quantity vs. shape", "Quantity vs. color", "Quantity vs. surface area")))%>%
  mutate(sample_quantity = factor(sample_quantity),
         feature = ifelse(experiment == "Quantity vs. surface area", feature_of_numerical_match, sample_feature),
         feature = ifelse(experiment == "Quantity only", "-",feature))%>%
  group_by(sample_quantity,feature, experiment)%>%
  tidyboot_mean(col = correct)


cols = c("#009E73" , "#E69F00","#56B4E9","#F0E442","#0072B2", "#D55E00","#000000")

ggplot()+
  geom_hline(yintercept = 0.5, lty = 2, alpha = .5)+
  geom_point(data = p3, aes(x = sample_quantity, y = mean, col = feature),position = position_dodge(width = .5), alpha = .25)+
  geom_pointrange(data = p4, aes(x = sample_quantity, y = mean, ymin = ci_lower, ymax = ci_upper, col = feature),position = position_dodge(width = .5), pch = 4)+
  geom_pointrange(data = p2, aes(x = sample_quantity, y = mean, ymin = ci_lower, ymax = ci_upper), col = "firebrick")+
  facet_grid(~experiment)+
  labs(y = "Proportion of choices based on quantity", x = "Quantity in training")+
  scale_color_manual(breaks=c("L", "triangle", "blue", "yellow", "double_size", "half_size"), name = "Feature", values = cols, na.value = "black")+
  theme_minimal()+
  theme(panel.border = element_rect(color = "black",fill = NA,size = 1))
  
ggsave("../figures/figure2.pdf", width = 10, height = 4)    
```

# Comparison to chance by experiment

```{r}
data %>%
  group_by(subject,experiment) %>%
  summarise(corr = mean(correct)) %>%
  group_by(experiment)%>%
  summarise(corr = list(corr))%>%
  group_by(experiment) %>%
  mutate(m = mean(unlist(corr)),
         sd = sd(unlist(corr)),
         df= t.test(unlist(corr), mu = 0.5)$parameter,
         t_value = t.test(unlist(corr), mu = 0.5)$statistic,
         p = t.test(unlist(corr), mu = 0.5)$p.value,
         d = cohensD(unlist(corr), mu = 0.5))%>%
  select(experiment,m,sd,t_value,df,p,d) 
```

Performance is above chance in `number_only` and `number_shape` and not different from chance in the other two conditions.

# Models by experiment

```{r}
model_data <-  data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface")), 
         sample_quantity = factor(sample_quantity)) %>%
  group_by(experiment)%>%
  mutate(block_individual = scale(block_individual), 
         trial_within_block = scale(trial_within_block))

```


## Number only

```{r}
mno <- glmer(correct ~ sample_quantity + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_only"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mno)

drop1(mno, test = "Chisq")
```

Significant effect of `sample_quantity`. Performance is pretty much at chance when they are trained on `1` (proportion correct = 0.47) and significantly higher (and probably also above chance) when trained on `4` (proportion correct = 0.67).

## Number vs. shape

```{r}
mnsh <- glmer(correct ~ sample_quantity + sample_feature + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_shape"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnsh)

drop1(mnsh, test = "Chisq")
```

No effect of `sample_feature`. Significant effect of `sample_quantity`. Performance is pretty much at chance when they are trained on `1` (proportion correct = 0.54) and significantly higher (and probably also above chance) when trained on `4` (proportion correct = 0.66).

## Number vs. color

```{r}
mnc <- glmer(correct ~ sample_quantity + sample_feature + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_color"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnc)

drop1(mnc, test = "Chisq")

```

No effect of `sample_quantity`. Significant effect of `sample_feature`. Performance is below chance when they are trained on `blue` (proportion correct = 0.36) and significantly higher (but not above chance) when trained on `yellow` (proportion correct = 0.52).

## Number vs. surface

```{r}
mnsu <- glmer(correct ~ sample_quantity + feature_of_numerical_match + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_surface"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnsu)

drop1(mnsu, test = "Chisq")

data %>%
  group_by(experiment)%>%
  summarise(n= length(unique(subject)))
  filter(experiment == "number_surface")%>%
  group_by(feature_of_numerical_match) %>%
  summarise(corr = mean(correct))
```

The sample in training was the same size in all cases. Therefore, we include the size fo the target at test (half size or double size) as a predictor. No effect of `sample_quantity` and no effect of `feature_of_numerical_match`.

# Model across experiemnts

```{r}
mo <- glmer(correct ~ experiment + sample_quantity + (block_individual + trial_within_block| subject),
      data = model_data,
      family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(mo)

drop1(mo, test = "Chisq")

emmeans(mo, pairwise ~ experiment, type = "response")$contrasts
```

Significant effects of `experiment` and `sample_quantity`. Pairwise contrasts show that performance does not differ between `number_only` and `number_shape` but performance in both experiments is significantly better than in `number_color` and `number_surface`, which do not differ from one another.     

