---
title: "Quantity judgements in stingless bees"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(tidyboot)
library(lme4)
library(ggthemes)
library(ggridges)
library(lsr)
library(broom.mixed)
library(emmeans)

```

# Data

```{r}
data <- read_csv("../data/data.csv")%>%
  mutate(subject = as.factor(subject))

```

# Sample size

```{r}
data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface")))%>%
  group_by(experiment)%>%
  summarise(N = length(unique(subject)))
```

# Visualize data by experiment

```{r}
p1 <- data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface"), labels  = c("Quantity only", "Quantity vs. shape", "Quantity vs. color", "Quantity vs. surface area")))%>%
  group_by(subject,experiment)%>%
  summarise(mean = mean(correct))#

p2 <- data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface"), labels  = c("Quantity only", "Quantity vs. shape", "Quantity vs. color", "Quantity vs. surface area")))%>%
  group_by(experiment)%>%
  tidyboot_mean(col = correct)

ggplot()+
  geom_hline(yintercept = 0.5, lty = 2, alpha = .5)+
  geom_violin(data = p1, aes(x = experiment, y = mean), alpha = .25, fill = "grey")+
  geom_point(data = p1, aes(x = experiment, y = mean), alpha = .5, col = "darkgrey")+
  geom_pointrange(data = p2, aes(x = experiment, y = mean, ymin = ci_lower, ymax = ci_upper), col = "firebrick")+
  ylim(0,1)+
  labs(y = "Proportion of choices based on quantity", x = "Experiment")+
  guides(col = F, fill = F)+
  scale_color_colorblind()+
  theme_minimal()
    
```
# Visualize data by experiment, sample quantity and feature (and overall)

```{r}

p2 <- p2 %>%
  mutate(sample_quantity = "overall",
         feature = "-")

p3 <- data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface"), labels  = c("Quantity only", "Quantity vs. shape", "Quantity vs. color", "Quantity vs. surface area")))%>%
  mutate(sample_quantity = factor(sample_quantity),
         feature = ifelse(experiment == "Quantity vs. surface area", feature_of_numerical_match, sample_feature),
         feature = ifelse(experiment == "Quantity only", "-",feature))%>%
  group_by(subject,sample_quantity,feature, experiment)%>%
  summarise(mean = mean(correct))#


p4 <- data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface"), labels  = c("Quantity only", "Quantity vs. shape", "Quantity vs. color", "Quantity vs. surface area")))%>%
  mutate(sample_quantity = factor(sample_quantity),
         feature = ifelse(experiment == "Quantity vs. surface area", feature_of_numerical_match, sample_feature),
         feature = ifelse(experiment == "Quantity only", "-",feature))%>%
  group_by(sample_quantity,feature, experiment)%>%
  tidyboot_mean(col = correct)


cols = c("#009E73" , "#E69F00","#56B4E9","#F0E442","#0072B2", "#D55E00","#000000")

ggplot()+
  geom_hline(yintercept = 0.5, lty = 2, alpha = .5)+
  geom_point(data = p3, aes(x = sample_quantity, y = mean, col = feature),position = position_dodge(width = .5), alpha = .25)+
  geom_pointrange(data = p4, aes(x = sample_quantity, y = mean, ymin = ci_lower, ymax = ci_upper, col = feature),position = position_dodge(width = .5), pch = 4)+
  geom_pointrange(data = p2, aes(x = sample_quantity, y = mean, ymin = ci_lower, ymax = ci_upper), col = "firebrick")+
  facet_grid(~experiment)+
  labs(y = "Proportion of choices based on quantity", x = "Quantity in training")+
  scale_color_manual(breaks=c("L", "triangle", "blue", "yellow", "double_size", "half_size"), name = "Feature", values = cols, na.value = "black")+
  theme_minimal()+
  theme(panel.border = element_rect(color = "black",fill = NA,size = 1))
  
ggsave("../figures/figure2.pdf", width = 10, height = 4)    
```

# Comparison to chance by experiment

```{r}
data %>%
  group_by(subject,experiment) %>%
  summarise(corr = mean(correct)) %>%
  group_by(experiment)%>%
  summarise(corr = list(corr))%>%
  group_by(experiment) %>%
  mutate(m = mean(unlist(corr)),
         sd = sd(unlist(corr)),
         df= t.test(unlist(corr), mu = 0.5)$parameter,
         t_value = t.test(unlist(corr), mu = 0.5)$statistic,
         p = t.test(unlist(corr), mu = 0.5)$p.value,
         d = cohensD(unlist(corr), mu = 0.5))%>%
  select(experiment,m,sd,t_value,df,p,d) 
```

Performance is above chance in `number_only` and `number_shape` and not different from chance in the other two conditions.

# Models by experiment

```{r}
model_data <-  data%>%
  mutate(experiment = factor(experiment, levels = c("number_only", "number_shape","number_color","number_surface")), 
         sample_quantity = factor(sample_quantity)) %>%
  group_by(experiment)%>%
  mutate(block_individual = scale(block_individual), 
         trial_within_block = scale(trial_within_block))

```


## Number only

```{r}
mno <- glmer(correct ~ sample_quantity + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_only"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mno)

drop1(mno, test = "Chisq")
```

Significant effect of `sample_quantity`. Performance is pretty much at chance when they are trained on `1` (proportion correct = 0.47) and significantly higher (and probably also above chance) when trained on `4` (proportion correct = 0.67).

## Number vs. shape

```{r}
mnsh <- glmer(correct ~ sample_quantity + sample_feature + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_shape"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnsh)

drop1(mnsh, test = "Chisq")
```

No effect of `sample_feature`. Significant effect of `sample_quantity`. Performance is pretty much at chance when they are trained on `1` (proportion correct = 0.54) and significantly higher (and probably also above chance) when trained on `4` (proportion correct = 0.66).

## Number vs. color

```{r}
mnc <- glmer(correct ~ sample_quantity + sample_feature + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_color"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnc)

drop1(mnc, test = "Chisq")

```

No effect of `sample_quantity`. Significant effect of `sample_feature`. Performance is below chance when they are trained on `blue` (proportion correct = 0.36) and significantly higher (but not above chance) when trained on `yellow` (proportion correct = 0.52).

## Number vs. surface

```{r}
mnsu <- glmer(correct ~ sample_quantity + feature_of_numerical_match + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_surface"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnsu)

drop1(mnsu, test = "Chisq")
```

The sample in training was the same size in all cases. Therefore, we include the size of the target at test (half size or double size) as a predictor. No effect of `sample_quantity` and no effect of `feature_of_numerical_match`.

# Model across experiemnts

```{r}
mo <- glmer(correct ~ experiment + sample_quantity + (block_individual + trial_within_block| subject),
      data = model_data,
      family = binomial,control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

summary(mo)

drop1(mo, test = "Chisq")

emmeans(mo, pairwise ~ experiment, type = "response")$contrasts
```

Significant effects of `experiment` and `sample_quantity`. Pairwise contrasts show that performance does not differ between `number_only` and `number_shape` but performance in both experiments is significantly better than in `number_color` and `number_surface`, which do not differ from one another.     

# Appendix: Interaction models

The structure of the data in the number vs shape, number vs. color, and number vs. surface experiments allows for modelling an interaction between the quantity of the sample and the feature of the sample (or the numerical match). In fact, when modelling this interaction, it comes out significant for all models. However, we have no good explanation for these interactions and we are unsure if they reflect a meaningful pattern or are an idiosyncrasy of the data that would unlikely replicate using slightly modified stimuli. We therefore do not discuss them in detail in the manuscript. However, for completeness, we want to include these models here. 

## Number vs. shape

```{r}
mnshi <- glmer(correct ~ sample_quantity * sample_feature + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_shape"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnshi)

drop1(mnshi, test = "Chisq")
```

Significant interaction between `sample_quantity` and no effect of `sample_feature`. Performance was highest when the sample feature was L-shaped and the sample quantity was 4. 

## Number vs. color

```{r}
mnci <- glmer(correct ~ sample_quantity * sample_feature + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_color"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnci)

drop1(mnci, test = "Chisq")

```

Significant interaction between `sample_quantity` and `sample_feature`. Performance was lowest when the sample quantity was 1 and the color was blue.   

## Number vs. surface

```{r}
mnsui <- glmer(correct ~ sample_quantity * feature_of_numerical_match + (block_individual + trial_within_block|subject), 
        data = model_data %>%filter(experiment == "number_surface"),
        family = binomial,
        control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5))
        )

summary(mnsui)

drop1(mnsui, test = "Chisq")
```

Significant interaction between `sample_quantity` and `feature_of_numerical_match`. When sample quantity was 1, performance was higher when the surface size was double, when the sample quantity was 4, this pattern was reversed. 